---
output:
  md_document:
    variant: markdown_github
    includes:
      in_header: header.md
    toc: true
    toc_depth: 2
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 4)
```

# Setup

## Software Libraries

## Mac OS X

```bash
## To install Homebrew:
/usr/bin/ruby -e "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/master/install)"
## Then:
brew tap homebrew/versions && brew tap homebrew/science && brew update
# brew install gcc --enable-cxx && brew link --overwrite gcc && brew link cmake
brew install boost --c++11
# Installing cmake may require: sudo chown -R `whoami` /usr/local/share/man/man7
brew install mlpack
```

## Ubuntu/Debian

```bash
sudo apt-get install libmlpack
```

## R Packages

```R
install.packages(c("BH", "Rcpp", "RcppArmadillo", "microbenchmark"))
```

If you get "ld: library not found for -lgfortran" error when trying to install RcppArmadillo, run:

```bash
curl -O http://r.research.att.com/libs/gfortran-4.8.2-darwin13.tar.bz2
sudo tar fvxz gfortran-4.8.2-darwin13.tar.bz2 -C /
```

See "[Rcpp, RcppArmadillo and OS X Mavericks "-lgfortran" and "-lquadmath" error](http://thecoatlessprofessor.com/programming/rcpp-rcpparmadillo-and-os-x-mavericks-lgfortran-and-lquadmath-error/)"" for more info.

# Rcpp

See [this section](http://rmarkdown.rstudio.com/authoring_knitr_engines.html#rcpp) in [RMarkdown documentation](http://rmarkdown.rstudio.com/) for details on Rcpp chunks.

```{r load_pkg}
library(magrittr)
library(Rcpp)
library(RcppArmadillo)
library(microbenchmark)
```

## Basics

```{Rcpp x2_source, cache = TRUE}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector cumSum(NumericVector x) {
  int n = x.size();
  NumericVector out(n);
  out[0] = x[0];
  for (int i = 1; i < n; ++i) {
    out[i] = out[i-1] + x[i];
  }
  return out;
}
```

```{r x2_example, dependson = 'x2_source', cache = TRUE}
x <- 1:1000
microbenchmark(
  native = cumsum(x),
  loop = (function(x) {
    output <- numeric(length(x))
    output[1] <- x[1]
    for (i in 2:length(x)) {
      output[i] <- output[i-1] + x[i]
    }
    return(output)
  })(x),
  Rcpp = cumSum(x)
) %>% summary(unit = "ms") %>% knitr::kable(format = "markdown")
```

## Using Libraries

This is the example code taken from Dirk's talk on RcppArmadillo and I'm using it to test how much I can do within Rcpp chunks in RMarkdown.

```{Rcpp fastlm_source, cache = TRUE}
// [[Rcpp::depends(RcppArmadillo)]]
#include <RcppArmadillo.h>
using namespace Rcpp;
using namespace arma;
// [[Rcpp::export]]
List fastLm(const colvec& y, const mat& X) {
  int n = X.n_rows, k = X.n_cols;
  colvec coef = solve(X, y);
  colvec resid = y - X*coef;
  double sig2 = as_scalar(trans(resid) * resid/(n-k));
  colvec stderrest = sqrt(sig2 * diagvec( inv(trans(X)*X)) );
  return List::create(_["coefficients"] = coef,
                      _["stderr"]       = stderrest,
                      _["df.residual"]  = n - k );
}
```

```{r fastlm_example, dependson = 'fastlm_source', cache = TRUE}
data("mtcars", package = "datasets")
microbenchmark(
  lm = lm(mpg ~ wt + disp + cyl + hp, data = mtcars),
  fastLm = fastLm(mtcars$mpg, cbind(1, as.matrix(mtcars[, c("wt", "disp", "cyl", "hp")]))),
  RcppArm = fastLmPure(cbind(1, as.matrix(mtcars[, c("wt", "disp", "cyl", "hp")])), mtcars$mpg)
) %>% summary(unit = "ms") %>% knitr::kable(format = "markdown")
```

Okay, let's try to adapt [RcppMLPACK's example code](https://github.com/thirdwing/RcppMLPACK/wiki/Example#k-means-example) for [k-means](http://www.mlpack.org/docs/mlpack-2.0.3/doxygen.php?doc=kmtutorial.html#kmeans_kmtut) to a RcppMLPACK-less context :)

```{Rcpp kmeans_source, eval = FALSE}
// [[Rcpp::plugins(cpp11)]]
// [[Rcpp::depends(RcppArmadillo)]]

#include <RcppArmadillo.h>
using namespace Rcpp;
#include <mlpack/methods/kmeans/kmeans.hpp>
using namespace mlpack::kmeans;

// [[Rcpp::export]]
arma::Row<size_t> kMeans(const arma::mat& data, const int& clusters) {
    arma::Row<size_t> assignments;
    KMeans<> k;
    k.Cluster(data, clusters, assignments); 
    return assignments;
}
```

I get the following:

```
Building shared library for Rcpp code chunk...

Quitting from lines 132-147 (README.Rmd) 
Error in dyn.load("/Users/mpopov/Desktop/Learning Rcpp/README_cache/markdown_github/kmeans_source_sourceCpp/sourceCpp-x86_64-apple-darwin13.4.0-0.12.7/sourcecpp_450b02624e/sourceCpp_2.so") : 
  unable to load shared object '/Users/mpopov/Desktop/Learning Rcpp/README_cache/markdown_github/kmeans_source_sourceCpp/sourceCpp-x86_64-apple-darwin13.4.0-0.12.7/sourcecpp_450b02624e/sourceCpp_2.so':
  dlopen(/Users/mpopov/Desktop/Learning Rcpp/README_cache/markdown_github/kmeans_source_sourceCpp/sourceCpp-x86_64-apple-darwin13.4.0-0.12.7/sourcecpp_450b02624e/sourceCpp_2.so, 6): Symbol not found: __ZN6mlpack3Log4InfoE
  Referenced from: /Users/mpopov/Desktop/Learning Rcpp/README_cache/markdown_github/kmeans_source_sourceCpp/sourceCpp-x86_64-apple-darwin13.4.0-0.12.7/sourcecpp_450b02624e/sourceCpp_2.so
  Expected in: flat namespace
 in /Users/mpopov/Desktop/Learning Rcpp/README_cache/markdown_github/kmeans_source_sourceCpp/sourceCpp-x86_64-apple-darwin13.4.0-0.12.7/sourcecpp_450b02624e/sourceCpp_2.so
Calls: <Anonymous> ... <Anonymous> -> source -> withVisible -> eval -> eval -> dyn.load
Execution halted
```

```{r kmeans_example, eval = FALSE}
data(trees, package = "datasets")
microbenchmark(
  kmeans = kmeans(trees, 3),
  kMeans = kMeans(t(trees), 3)
) %>% summary(unit = "ms") %>% knitr::kable(format = "markdown")
```

## Modules

### Simple

# References

- Eddelbuettel, D. (2013). Seamless R and C++ Integration with Rcpp. New York, NY: Springer Science & Business Media. http://doi.org/10.1007/978-1-4614-6868-4
